MCMC convergence diagnostics

### 2023-12-05 src/naomi-simple_mcmc/mcmc-convergence.Rmd
Final improvements to NUTS figures

(rhat_min <- min(mcmc_summary[, "Rhat"]))
(rhat_lower <- quantile(mcmc_summary[, "Rhat"], 0.025))
(rhat_median <- quantile(mcmc_summary[, "Rhat"], 0.50))
(rhat_upper <- quantile(mcmc_summary[, "Rhat"], 0.975))
(rhat_max <- max(mcmc_summary[, "Rhat"]))
(rhat_mean <- mean(mcmc_summary[, "Rhat"]))
```

@@

ess_fig <- bayesplot::mcmc_neff_data(ratios) %>%
  mutate(
    description_new = case_when(
      value < 0.1 ~ "< 0.1",
      value < 0.5 ~ "< 0.5",
      value > 0.5 ~ "> 0.5"
    )
  ) %>%
  ggplot(mapping = aes(x = value, y = parameter, color = description_new)) +
  geom_segment(aes(yend = parameter, xend = -Inf), na.rm = TRUE, alpha = 0.7) +
  scale_color_manual(values = c("#E69F00", "#009E73", "#56B4E9")) +
  geom_vline(xintercept = 0.1, linetype = "dashed", col = "grey40", alpha = 0.5) +
  geom_vline(xintercept = 0.5, linetype = "dashed", col = "grey40", alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", col = "grey40", alpha = 0.5) +
  labs(x = "ESS ratio", y = "Parameter", col = "", tag = "A") +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )

@@

```{r}
ess_histogram <- data.frame(mcmc_summary) %>%
  tibble::rownames_to_column("param") %>%
  ggplot(aes(x = n_eff)) +
    geom_histogram(alpha = 0.8) +
    labs(x = "ESS", y = "Count", tag = "B")

@@

```{r}
out <- list(
  "rhat_min" = rhat_min,
  "rhat_lower" = rhat_lower,
  "rhat_median" = rhat_median,
  "rhat_upper" = rhat_upper,
  "rhat_max" = rhat_max,
  "rhat_mean" = rhat_mean,
  "ess_min" = ess_min,
  "ess_lower" = ess_lower,
  "ess_median" = ess_median,
  "ess_upper" = ess_upper,
  "ess_max" = ess_max,
  "ess_mean" = ess_mean
)

@@

```{r}
(plot <- bayesplot::mcmc_trace(mcmc, pars = c(names(which.min(mcmc_summary[, "n_eff"])), names(which.max(rhats)))))

ggsave("worst-trace.png", plot, h = 3, w = 6.25)
```

>>>>

(rhat_min <- min(mcmc_summary[, "Rhat"]))
(rhat_lower <- quantile(mcmc_summary[, "Rhat"], 0.025))
(rhat_median <- quantile(mcmc_summary[, "Rhat"], 0.50))
(rhat_upper <- quantile(mcmc_summary[, "Rhat"], 0.975))
(rhat_max <- max(mcmc_summary[, "Rhat"]))
(rhat_mean <- mean(mcmc_summary[, "Rhat"]))
(rhat_above_threshold <- mean(mcmc_summary[, "Rhat"] < 1.01))
```

@@

ess_fig <- bayesplot::mcmc_neff_data(ratios) %>%
  mutate(
    description_new = case_when(
      value < 0.1 ~ "< 0.1",
      value < 0.5 ~ "< 0.5",
      value > 0.5 ~ "> 0.5"
    )
  ) %>%
  ggplot(mapping = aes(x = value, y = parameter, color = description_new)) +
  geom_segment(aes(yend = parameter, xend = -Inf), na.rm = TRUE, alpha = 0.7) +
  scale_color_manual(values = c("#E69F00", "#009E73", "#56B4E9")) +
  geom_vline(xintercept = 0.1, linetype = "dashed", col = "grey40", alpha = 0.5) +
  geom_vline(xintercept = 0.5, linetype = "dashed", col = "grey40", alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", col = "grey40", alpha = 0.5) +
  labs(x = "Effective sample size ratio", y = "Parameter", col = "", tag = "A") +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )

@@

```{r}
ess_histogram <- data.frame(mcmc_summary) %>%
  tibble::rownames_to_column("param") %>%
  ggplot(aes(x = n_eff)) +
    geom_histogram(col = "grey60", fill = "grey80") +
    labs(x = "Effective sample size", y = "Count", tag = "B")

@@

```{r}
out <- list(
  "rhat_min" = rhat_min,
  "rhat_lower" = rhat_lower,
  "rhat_median" = rhat_median,
  "rhat_upper" = rhat_upper,
  "rhat_max" = rhat_max,
  "rhat_mean" = rhat_mean,
  "rhat_above_threshold" = rhat_above_threshold,
  "ess_min" = ess_min,
  "ess_lower" = ess_lower,
  "ess_median" = ess_median,
  "ess_upper" = ess_upper,
  "ess_max" = ess_max,
  "ess_mean" = ess_mean
)

@@

```{r}
worst_eff <- bayesplot::mcmc_trace(mcmc, pars = names(which.min(mcmc_summary[, "n_eff"]))) + labs(tag = "A") + guides(col = "none")
worst_rhat <- bayesplot::mcmc_trace(mcmc, pars = names(which.max(rhats))) + labs(tag = "B")

worst_eff + worst_rhat

ggsave("worst-trace.png", h = 3, w = 6.25)
```
